Simulations of fold number selection in cross-validation of linear and LASSO regression

### 2023-09-02 manuscript/manuscript.Rmd
“Sept2”

---
title: "Simulations of fold number selection in cross-validation of linear and LASSO regression"
author:
- Angelos Vasilopoulos
- Gregory J. Matthews
abstract: "*k*-fold cross-validation is a popular method of error estimation for model selection in computational research. However, there is limited focus in the literature on the question of what fold number *k* is appropriate for various dataset dimensions. Here we review relevant literature and present a simulation of linear and least absolute shrinkage and selection operator (LASSO) regression prediction error estimation at various values of *k* and sample size *n*. In agreement with a growing body of literature, we find that contrary to a persisting understanding, there is no bias-variance trade-off in selection of *k*. Instead, with increasing *k* both bias and variance decrease, perhaps asymptotically. Our results also suggest a predictable relationship between optimal values of *k* and *n*. | *Keywords*: cross-validation, optimization, fold number"
output:
  pdf_document:
    df_print: kable
    number_sections: yes
    keep_tex: yes
  word_document: default
fontsize: 12pt
link-citations: yes
linkcolor: cyan
urlcolor: cyan
bibliography: /Users/angelos/Documents/research/optimal_k/manuscript/optimal_k_references.bib
header-includes: \usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t} \usepackage[labelsep=period]{caption}
indent: TRUE
---

@@

>>>>

---
title: "Simulations of fold number selection in cross-validation of linear and LASSO regression"
author:
- Angelos Vasilopoulos
- Gregory J. Matthews
abstract: "*k*-fold cross-validation is a popular method of error estimation for model selection in computational research. However, there is limited focus in the literature on the question of what fold number *k* is appropriate for various dataset dimensions. Here we review relevant literature and present a simulation of linear and least absolute shrinkage and selection operator (LASSO) regression prediction error estimation at various values of *k* and sample size *n*. In agreement with a growing body of literature, we find that contrary to a persisting understanding, there is no bias-variance trade-off in selection of *k*. Instead, with increasing *k* both bias and variance decrease, perhaps asymptotically. Our results also suggest a predictable relationship between optimal values of *k* and *n*. | *Keywords*: cross-validation, optimization, fold number"
output:
  pdf_document:
    df_print: kable
    number_sections: yes
    keep_tex: yes
  word_document: default
fontsize: 12pt
link-citations: yes
linkcolor: cyan
urlcolor: cyan
bibliography: /Users/angelos/optimal_k_git/manuscript/optimal_k_references.bib
header-includes: \usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t} \usepackage[labelsep=period]{caption}
indent: TRUE
---

@@

Note that the case of linear regression with $n = 100$ and number of features $p = 100$ is not displayed because there are $n - p - 1 = -1$ degrees of freedom, which is not a sufficient number for regression. The case of $n = 100$ and $p = 95$ degrees of freedom is similarly problematic and is not displayed because the high level of error in bias, variance, and $\mathrm{MSE}$ estimation associated with $n - p - 1 = 5$ degrees of freedom obscures important patterns in the rest of the data.

### 2023-09-02 manuscript/manuscript.tex
“Sept2”

>>>>

Note that the case of linear regression with \(n = 100\) and number of
features \(p = 100\) is not displayed because there are
\(n - p - 1 = -1\) degrees of freedom, which is not a sufficient number
for regression. The case of \(n = 100\) and \(p = 95\) degrees of
freedom is similarly problematic and is not displayed because the high
level of error in bias, variance, and \(\mathrm{MSE}\) estimation
associated with \(n - p - 1 = 5\) degrees of freedom obscures important
patterns in the rest of the data.
